Graph Databases for Use with Timeseries IoT Datasets

### 2023-11-29 paper.tex
Edits to Final Paper
Made many edits to final paper

It was decided that using a graph database could provide answers to these questions
more efficiently and easily than a relational database. The primary advantage for use in this paper
is the fact that graph databases are great for use with densely connected data \cite{GraphSurvey}.
On top of this, they are very quick to query and will give a result relatively quickly compared to
other solutions \cite{GraphSurvey}.

\section{Related Works}
There are many current systems that implement different components of the general idea of graphical databases for
IoT device management andanomoly detection, but none put these components together. \textit{IoT Inspector} is a great
tool that performs the task of monitoring using a form of a relational database, but not a graphical one \cite{IoTInspector}.
The authors of \textit{The Graph of Things} created such a system for aggregating IoT devices worldwide \cite{GraphofThings}; 
However, these systems are not built for small networks or home users.

\section{Methodology}
The first step in this project was choosing the technology that was to be used. It was decided to use Neo4J
and Python for the tech stack.

@@

@@

@@

@@

@@

\section{Analysis}

>>>>

It was decided that using a graph database could provide answers to these questions
more efficiently and easily than a relational database. The primary advantage for use in this paper
is the fact that graph databases are great for use with densely connected data \cite{GraphSurvey}.
On top of this, they are very quick to query and will give a result relatively quickly compared to
other solutions \cite{GraphSurvey}. In particular, the use for home users and smaller networks was 
examined by asking what queries could be run on a smaller dataset as well as what would be reasonable
for a home user as far as time and efficiency.

\section{Related Works}
There are many current systems that implement different components of the general idea of graphical databases for
IoT device management andanomoly detection, but none put these components together. \textit{IoT Inspector} is a great
tool that performs the task of monitoring using a form of a relational database, but not a graphical one \cite{IoTInspector}.
The authors of \textit{The Graph of Things} created such a system for aggregating IoT devices worldwide \cite{GraphofThings}; 
However, these systems are not built for small networks or home users. What this paper seeks to explore is what questions
can a graph database answer for these smaller networks, home users, and generally users with smaller amounts of
IP addresses being used. The queries and dataset going forward were chosen with these thoughts in mind.

\section{Methodology}
This projects main goal was to see the about the ease of use for a graph database for answering certain questions.
It was hypothesied that fairly simple queries to the database would allow the author to gain large amounts of
information to be analyzed and from which many conclusions could be drawn.

The first towards determining this hypothesis was choosing the technology that was to be used. It was decided 
to use Neo4J and Python for the tech stack. Neo4J was chosen due to its unique feature of having practically 
unlimited properties attached to both the nodes and the edges. This was particularly useful when working with 
5-tuples where the nodes could hold the addresses and the edged could hold the ports, protocals, and a timestamp.
Python was chosen due to its ease of use, quick prototyping, and library support for Neo4J.

The next step was aquaring the data. In this regard, after a long time of testing and consideration, it was
determined that using a prebuilt dataset would be the best and most time efficient. A perfect dataset was
found with BoT IoT \cite{BoTIoT}. This dataset includes both simulated and real world IoT devices as well
as showing many attacks on these devices. Using such a dataset would allow for getting great amounts of
patterns for which queries could be written to detect. In addition, being timeseries based, it was perfect
for this paper.

With both the tech stack and dataset decided, the next step was to decide what types of queries should be
written to get the most out of the potentials of both the dataset and graph databases as a whole. Some basic
queries were written with some queries left to future work due to lack of knowledge and time on the part of 
the author. The basic queries included queries such as flows to and from an IP address to determine whether
a device was a sink, source, or equal with regards to flow direction, how many flows were sent/received per
second, how many ports were contacted, and the number of incoming flows to a particular address. These queries
were chosen in particular to look at two certain attacks: DDoS and Network Scan attacks. If an address were a 
sink, it could be a target, if it was a source it could be an attacker for DDoS. However to truly determine this
flows per second and incoming flows would also need to be detected. For Network Scans, if the number of ports
an IP address sends a flow to is abnormally high, it is a likely indicator that a network scan is being performed.

The last step was writting the program and aquiring the results. This produced some very interesting conclusions
and was satisfying to the author. This also happened to be the longest part of the methodology as parsing all the
data took a relatively long time, over a day and a half to parse 1,000,000 flows into the database. Once parsed,
however, the queries were very quick with the longest taking only a matter of seconds to execute.

@@

With the experiments done and the data aquired there were some interesing results that were discovered. One such
discovery was that, within the dataset being used, there were clear sinks and sources for flows as shown in figures
\ref{fig:flowfrom} and \ref{fig:flowto}.

@@

A clear pattern of sending and receiving was being established in the data which is a potential indicator of some
sort of attack whether DDoS, Network Scanning or otherwise. To determine this further, the average number of flows
per second was determined as shown in figures \ref{fig:sentfrompersec} and \ref{fig:senttopersec}.

@@

As with figures \ref{fig:flowfrom} and \ref{fig:flowto}, these figures also showed clear trends with definitive
sinks and sources, however this was not enough to fully determine what was going on with this data. To make a more
definitive conclusion the number of ports each IP address contacted was found as shown in figure \ref{fig:portscontacted}.

@@

This figure provided a more definitive idea about what was going on for some of the IP Addresses. In particular,
192.168.100.148 contacted the most ports likely indicating that it was performing a network scan on one of the
other networks. Similarily could be said for 192.168.100.149 and 192.168.100.150 with less confidence.

@@

\section{Discussion}
